---
title: "theme_usq"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Introduction

The goal of _theme.usq_ is to provide
[University of Southern Queensland](https://usq.edu.au) (USQ) staff and students
a quick and easy way to apply USQ colours and typography to graphs created in R
using _ggplot2_ while providing clear graphs for reports and presentations. All
of the colours provided are defined in USQ's Visual Identity Colour Palette, but
do not all appear in the same order to maintain usability for the purposes of
graphing.

This package has been tested on MacOS, Ubuntu Linux and USQ computers using
Windows 7. For Linux users, if you have not installed the MS Core Fonts, you
will need to do that for this package to function properly and generate the
graphs with the proper typography. Windows and MacOS users should be ready to go
with just the installation of this package.

# Quickstart

The _theme.usq_ package is only available from GitHub. The easiest way to
install it is by using the [_devtools_](https://github.com/hadley/devtools)
package.

The installation may take some time as some system fonts need to be catalogued
to use the Microsoft Verdana font that USQ suggests. Once the installation is
done, it should not be necessary to re-catalogue the fonts so loading
_theme.usq_ will not take any longer than expected.

```{r, eval=FALSE}
if(!require(ggplot2)){
  devtools::install.packages("ggplot2")
}

if(!require(devtools)){
    install.packages("devtools")
    library(devtools)
}

devtools::install_github("adamhsparks/theme_usq")
```

# Examples

## Example 1: Scatterplots of discrete data

Using the _usq.theme's_ `theme_usq` for _ggplot2_, plot car weights by
mpg and facet by `Transmission` (0 = automatic, 1 = manual) using the
`usq_palette` in the `scale_colour_manual` discrete scale function to use
USQ colours for the graph.

```{r}
library("ggplot2")
library("theme.usq")

# Discrete values
p1 <- ggplot(mtcars) +
  geom_point(aes(
    x = wt,
    y = mpg,
    colour = factor(gear)
  )) +
  scale_colour_manual(values = usq_palette) +
  facet_wrap(~ am)

p1
```

Now add the `theme_usq()` to the graph.

```{r}
p1 + theme_usq()
```

## Example 2: Heatmaps or other continuous data

Using the _usq.theme's_ `theme_usq()` for _ggplot2_, plot values using the
`usq_fill_gradient` to use USQ colours for continuous values in the graph. Two
types of gradients are included, warm and cool for both `scale_fill_gradient()`
and `scale_colour_gradient()` as necessary.

### Warm gradients

```{r}
p2a <- ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_raster(aes(fill = density), interpolate = TRUE) +
  usq_fill_gradient_warm() +
  theme_usq()
  
p2a
```

### Cool gradients

```{r}
p2b <- ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_raster(aes(fill = density), interpolate = TRUE) +
  usq_fill_gradient_cool() +
  theme_usq()
  
p2b
```

## Example 3: Heatmaps using other colour palettes

`theme_usq()` can be used with any colour palette that you wish to use, while
still applying the graph styling and typography to the graph.

Use the default _ggplot2_ colour scheme to fill the density plot while using
the `theme_usq()` to theme the graph.

```{r}
p3 <- ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_raster(aes(fill = density), interpolate = TRUE) +
  theme_usq()
  
p3
```

## Example 4: Histograms

Use a USQ colour to fill a histogram by using the second value in the
`usq_palette` (USQ Charcoal) as the fill value, a good colour for presentations
that are going to be projected on a screen.

```{r}
p4 <- ggplot(diamonds, aes(carat)) +
  geom_histogram(fill = usq_palette[1]) +
  theme_usq()

p4
```

## Example 5: Boxplots

Use the USQ colours to fill the boxplots, while using the first value in the
`usq_palette` (USQ Charcoal) as the colour value.

```{r}
p5 <- ggplot(mpg, aes(class, hwy)) +
  geom_boxplot(aes(fill = drv), colour = usq_palette[1]) +
  scale_fill_manual(values = usq_palette) +
  theme_usq()
  
p5
```

## Example 6: Timeseries

Plot timeseries lines using discrete colours for each variable of interest.

```{r}
p6 <- ggplot(economics_long, aes(date, value01, colour = variable)) +
  geom_line() +
  scale_colour_manual(values = usq_palette) +
  theme_usq()
  
p6
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
