---
title: "theme_usq"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/adamhsparks/theme_usq.svg?branch=master)](https://travis-ci.org/adamhsparks/theme_usq)

# theme\_usq

The goal of _theme\_usq_ is to provide USQ staff and students with an easy to use
_ggplot2_ theme that conforms to the USQ visual guidelines for typography and
colour.

## Quickstart

The _theme.usq_ package is only available from GitHub. The easiest way to
install it is by using the [_devtools_](https://github.com/hadley/devtools) package.

Currently the development version of [_ggplot2_](https://github.com/tidyverse/ggplot2) is required and also needs
to be installed from GitHub.

The installation may take some time as some system fonts need to be catalogued
to use the Microsoft Verdana font that USQ suggests. Once the installation is
done, it should not be necessary to recatalogue the fonts so loading _theme.usq_
will not take any longer than expected.

```{r, eval=FALSE}
if(!require(devtools)){
    install.packages("devtools")
    library(devtools)
}

devtools::install_github("tidyverse/ggplot2")
devtools::install_github("adamhsparks/theme_usq")
```

## Example 1: Graphing Discrete Data

Using the _theme.usq's_ `theme_usq` for _ggplot2_, plot car weights by
mpg and facet by `Transmission` (0 = automatic, 1 = manual) using the
`usq_palette` in the `scale_colour_manual` discrete scale function to use
USQ colours for the graph.

```{r}
library("ggplot2")
library("theme.usq")

# Discrete values
p1 <- ggplot(mtcars) +
  geom_point(aes(
    x = wt,
    y = mpg,
    colour = factor(gear)
  )) +
  scale_colour_manual(values = usq_palette) +
  facet_wrap(~ am) +
  theme_usq()
p1
```

## Example 2: Graphing Continuous Data

Using the _usq.theme's_ `theme_usq` for _ggplot2_, plot values using the
`usq_fill_gradient` to use USQ colours for continuous values in the graph.

```{r}
# continuous values
df <- reshape2::melt(outer(1:4, 1:4), varnames = c("X1", "X2"))

p2 <- ggplot(df, aes(X1, X2)) +
  geom_tile(aes(fill = value)) +
  usq_fill_gradient() +
  theme_usq()
p2
```


## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
