---
title: "theme_usq"
output: github_document
---

`
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/adamhsparks/theme_usq.svg?branch=master)](https://travis-ci.org/adamhsparks/theme_usq)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/adamhsparks/theme_usq?branch=master&svg=true)](https://ci.appveyor.com/project/adamhsparks/theme_usq)
[![Coverage Status](https://img.shields.io/codecov/c/github/adamhsparks/theme_usq/master.svg)](https://codecov.io/github/adamhsparks/theme_usq?branch=master)

# theme\_usq

The goal of _theme\_usq_ is to provide USQ staff and students with an easy to use
_ggplot2_ theme that conforms to the USQ visual guidelines for typography and
colour.


## Example 1: Graphing Discrete Data

Using the _usq.theme's_ `theme_usq` for _ggplot2_, plot car weights by
mpg and facet by `Transmission` (0 = automatic, 1 = manual) using the
`usq_palette` in the `scale_colour_manual` discrete scale function to use
USQ colours for the graph.

```{r}
library("ggplot2")
library("theme.usq")

# Discrete values
p1 <- ggplot(mtcars) +
  geom_point(aes(
    x = wt,
    y = mpg,
    colour = factor(gear)
  )) +
  scale_colour_manual(values = usq_palette) +
  facet_wrap(~ am) +
  theme_usq()
p1
```

## Example 2: Graphing Continuous Data

Using the _usq.theme's_ `theme_usq` for _ggplot2_, plot values using the
`usq_fill_gradient` to use USQ colours for continuous values in the graph.

```{r}
# continuous values
df <- reshape2::melt(outer(1:4, 1:4), varnames = c("X1", "X2"))

p2 <- ggplot(df, aes(X1, X2)) +
  geom_tile(aes(fill = value)) +
  usq_fill_gradient() +
  theme_usq()
p2
```


## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
