---
title: "theme.usq"
author: "Adam H Sparks"
date: "`r Sys.Date()`"
output:
  html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{theme.usq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, eval = TRUE, message=FALSE, echo=FALSE}
library(theme.usq)
library(ggplot2)
knitr::opts_chunk$set(eval = TRUE, collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5, fig.align = "center")
```

## Introduction to _theme.usq_

The goal of _**`theme.usq`**_ is to provide 
[University of Southern Queensland](https://usq.edu.au) (USQ) staff and students
a quick and easy way to apply USQ colours and typography to graphs created in R
using the base _**`graphics`**_ package or _**`ggplot2`**_ while providing clear graphs
for reports and presentations. All of the colours provided are defined in USQ's
Visual Identity Colour Palette, but do not all appear in the same order to
maintain usability for the purposes of graphing.

This package has been tested on macOS, Ubuntu Linux and USQ computers using
Windows 7. For Linux users, if you have not installed the MS Core Fonts, you
will need to do that for this package to function properly and generate the
graphs with the proper typography. Windows and macOS users should be ready to go
with just the installation of this package.

# Using _theme.usq_

Load the library using the following command in R.
```{r eval=FALSE}
library(theme.usq)
```

## Colours

To view a list of named colours in this package, simply type:
```{r list-colours, eval=TRUE}
usq_colours
```

The resulting list shows the hexadecimal colour, e.g. "#ffd100", and it's name,
"usq yellow". The names may be used to specify the colours at any time when
plotting.

## Palettes

Colours are grouped into convenient palettes as well. The palettes may be used
when plotting. To see the palettes and colours included in each of them, simply
type:

```{r list-palettes, eval=TRUE}
usq_palettes
```

## Examples

Following are a few examples of _**`theme.usq's`**_ capabilities.

### Example 1: Scatterplots of discrete data

#### Using plot_usq()

Plot plot car weights by miles per gallon.

```{r basic-plot}
plot_usq(x = mtcars$wt, y = mtcars$mpg)
```

#### Using _ggplot2_ and theme_usq()

Plot car weights by miles per gallon and facet by `Transmission` (0 = automatic,
1 = manual) using the `scale_colour_usq()` setting `discrete = TRUE` and using
the `cool` palette to use blue USQ colours for the plot.

```{r basic-ggplot2}
p1 <- ggplot(data = mtcars) +
  geom_point(aes(
    x = wt,
    y = mpg,
    colour = factor(gear)
  )) +
  scale_colour_usq(discrete = TRUE,
                   palette = "cool") +
  facet_wrap(~ am)

p1
```

Now add the `theme_usq()` to the graph.

```{r ggplot2-theme-usq}
p1 + theme_usq()
```

### Example 2: Heatmaps or other continuous data

Using the _**`theme.usq's`**_ `theme_usq()` for _**`ggplot2`**_, plot values
using the `scale_fill_usq()` to use USQ colours for continuous values in the
graph.

#### Warm gradients

```{r warm-gradients}
p2a <- ggplot(data = faithfuld, aes(x = waiting, y = eruptions)) +
  geom_raster(aes(fill = density),
              interpolate = TRUE) +
  scale_fill_usq("warm",
                 discrete = FALSE) +
  theme_usq()

p2a
```

#### Cool gradients

```{r cool-gradients}
p2b <- ggplot(data = faithfuld, aes(x = waiting, y = eruptions)) +
  geom_raster(aes(fill = density),
              interpolate = TRUE) +
  scale_fill_usq("cool",
                 discrete = FALSE) +
  theme_usq()
  
p2b
```

### Example 3: Heatmaps using other colour palettes

#### Using _ggplot2_ and theme_usq()

`theme_usq()` can be used with any colour palette that you wish to use, while
still applying the graph styling and typography to the graph.

Use the default _**`ggplot2`**_ colour scheme to fill the density plot while
using the `theme_usq()` to theme the graph.

```{r theme-usq-other-pals}
p3 <- ggplot(data = faithfuld, aes(x = waiting, y = eruptions)) +
  geom_raster(aes(fill = density), interpolate = TRUE) +
  theme_usq()
  
p3
```

### Example 4: Histograms

#### Using hist_usq()

```{r base-hist}
hist_usq(diamonds$carat)
```

#### Using _ggplot2_ and theme_usq()

```{r ggplot2_hist}
p4 <- ggplot(data = diamonds, aes(x = carat)) +
  geom_histogram(fill = usq_cols("usq charcoal")) +
  theme_usq()

p4
```

### Example 5: Boxplots

#### Using boxplot_usq()

Plot the highway miles per gallon (mpg) of 38 popular car models in the US by
class of car.

```{r base-boxplot}
boxplot_usq(mpg$hwy ~ mpg$class)
```

#### Using _ggplot2_ and theme_usq()

```{r ggplot2-boxplot}
p5 <- ggplot(data = mpg, aes(x = class, y = hwy)) +
  geom_boxplot(alpha = 0.75,
               fill = usq_cols("usq charcoal"),
               colour = usq_cols("usq charcoal")) +
  theme_usq()
  
p5
```

Use the USQ colours to fill the box-plots while using `drv` (_e.g._, 4-wheel
drive, front-wheel drive or rear-wheel drive) for the box-plot colour.

```{r ggplot2-boxplot-fill}
p5.1 <- ggplot(data = mpg, aes(x = class, y = hwy)) +
  geom_boxplot(aes(fill = drv),
               colour = usq_cols("usq charcoal")) +
  scale_fill_usq("primary") +
  theme_usq()
  
p5.1
```

### Example 6: Timeseries

#### Using _ggplot2_ and theme_usq() to plot timeseries lines using
discrete colours for each variable of interest. While possible to do with base R
graphics, _**`ggplot2`**_ simplifies the process greatly, so it is the only
example provided and suggested for use.

```{r timeseries-lines}
p6 <- ggplot(data = as.data.frame(economics_long),
             aes(x = date, y = value01, colour = variable)) +
  geom_line() +
  scale_colour_usq("bright") +
  theme_usq()
  
p6
```

### Example 7: Barplots

#### Using barplot_usq()

Plot the areas in thousands of square miles of landmasses which exceed 10,000
sqm.

```{r base-barplots}
barplot_usq(islands, col = "cool gray")
```

#### Using _ggplot2_ and theme_usq()

Plot the areas in thousands of square miles of landmasses which exceed 10,000
sqm. Note the use of `coord_flip()` to rotate the axis so that the x-axis labels
do not need to be rotated.

```{r ggplot2-barplot}
library(tibble)
islands_df <- as.data.frame(islands)
islands_df <- rownames_to_column(islands_df, "name")

ggplot(data = islands_df, aes(x = name, y = islands)) +
  geom_bar(stat = "identity", 
           colour = usq_cols("cool gray"),
           fill = usq_cols("cool gray")) +
  theme_usq() +
  coord_flip()
```

Plot the number of cars from each manufacturer in the `mpg()` dataset.

```{r example-bright}
ggplot(data = mpg, aes(x = manufacturer, fill = manufacturer)) +
  geom_bar() +
  scale_fill_usq(palette = "bright", guide = "none") +
  theme_usq() +
  coord_flip()
```

## Using theme_usq() in presentations

The default settings for `theme_usq()` are fine for printed materials such as
reports, but if you wish to use it in a presentation you can use the
_**`ggplot2`**_ option `base_size = ` to increase the size of the fonts, points
and lines in the graphs.

### Example 8: Using _ggplot2_'s base_size() with theme_usq()

As an example, using `base_size = 24` helps ensure that the graphs are legible
on a standard sized PowerPoint slide when exported at a size of 33.87mm x
19.05mm, using `ggsave()`, for a 16x9 presentation slide that fills the whole
slide.

```{r ggplot2-presentation, eval = FALSE}
p8 <-  ggplot(data = mpg, aes(x = class, y = hwy)) +
  geom_boxplot(fill = usq_cols("usq charcoal"),
               colour = usq_palette("usq charcoal"),
               alpha = 0.75) +
  theme_usq(base_size = 24) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("slide_p7.png", width = 33.87, height = 19.05, units = "cm",
       dpi = 150)
```
