---
title: "theme_usq"
author: "Adam H Sparks"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{theme_usq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Example 1: Graphing Discrete Data

Using the _usq.theme's_ `theme_usq` for _ggplot2_, plot car weights by
mpg and facet by `Transmission` (0 = automatic, 1 = manual) using the
`usq_palette` in the `scale_colour_manual` discrete scale function to use
USQ colours for the graph.

```{r}
library("ggplot2")
library("theme.usq")

# Discrete values
p1 <- ggplot(mtcars) +
  geom_point(aes(
    x = wt,
    y = mpg,
    colour = factor(gear)
  )) +
  scale_colour_manual(values = usq_palette) +
  facet_wrap(~ am) +
  theme_usq()
p1
```

## Example 2: Graphing Continuous Data

Using the _usq.theme's_ `theme_usq` for _ggplot2_, plot values using the
`usq_fill_gradient` to use USQ colours for continuous values in the graph.

```{r}
# continuous values
df <- reshape2::melt(outer(1:4, 1:4), varnames = c("X1", "X2"))

p2 <- ggplot(df, aes(X1, X2)) +
  geom_tile(aes(fill = value)) +
  usq_fill_gradient() +
  theme_usq()
p2
```
